---
title: "Stokes cut comparer"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
runtime: shiny
resource_files:
- out/joined.rds
- out/cagrs.rds
- out/internal_vs_stokes.rds
- data/lmo64_agg_stokes_mapping.csv
- out/lfs_data.csv
---

```{r, include=FALSE}
library(tidyverse)
library(here)
library(readxl)
library(janitor)
library(assertthat)
library(conflicted)
library(htmlwidgets)
conflicts_prefer(dplyr::filter)
options(scipen = 999)
#functions-----------------------
cagr_plot <- function(tbbl, xvar, yvar){
  plt <- ggplot(cagrs, aes({{  xvar  }}, {{  yvar  }},
                         text=paste0("Industry: ",
                                     industry,
                                     "\n Internal CAGR = ",
                                     scales::percent({{  xvar  }}, accuracy = .1),
                                     "\n Stokes CAGR = ",
                                     scales::percent({{  yvar  }}, accuracy = .1))))+
  geom_abline(slope=1, intercept = 0, colour="white", lwd=1.5)+
  geom_point(alpha=.5)+
  scale_x_continuous(labels=scales::percent)+
  scale_y_continuous(labels=scales::percent)+
  labs(x="internal forecast",
       y="stokes forecast")
  plotly::ggplotly(plt, tooltip="text")
}
```

```{r}
joined <- read_rds(here("out","joined.rds"))
internal_vs_stokes <- read_rds(here("out","internal_vs_stokes.rds"))
cagrs <- read_rds(here("out","cagrs.rds"))
mapping <- read_csv(here("data","lmo64_agg_stokes_mapping.csv"))
lfs_data <- read_csv(here("out","lfs_data.csv"))
files <- sort(unique(joined$which_file), decreasing = TRUE)
```

Comparison with previous cut
=====================================  

Inputs {.sidebar data-width=400}
-------------------------------------

*  This app allows one to visualize what has changed between Stokes cuts, and how employment compares with our internal forecast.
*  Here we are looking at the difference between the 4th and 5th Macro cut of 2023.
*  Note that the data series (rows) differ in several orders of magnitude in terms of their minimum absolute value.
*  For series with a large $\min(|original~value|)$, we calculate the percentage change between the cuts:
$$change=\frac{new~value-original~value}{|original~value|}\times100$$
*  For series with a small $min(|original~value|)$, we subtract the original value from the new value:
$$change = new~value - original~value$$
*  Doing so puts all series on a similar scale, allowing us to use a heatmap to identify the significant changes between cuts.
*  You can hover over the cells of the heatmap to get more details.


```{r}
selectInput(
  "file",
  "Select an excel file to compare",
  files,
  selected = files[1]
)

filtered <- reactive({
  joined|>
    filter(which_file==input$file)
})

renderUI({#need the render UI because the list of choices depends on the excel file
selectInput(
  "sheet",
  "Select a sheet to compare",
  unique(filtered()$sheet),
  selected = unique(filtered()$sheet)[1]
)
})

# selectInput(
#   "year",
#   "Only consider data since",
#   2013:2023,
#   2013
# )

numericInput(
  "minimum",
  "min(|original value|) threshold",
  10,
  min = 1,
  max = 100,
  step = 1
)

reactive_tbbl <- reactive({
  temp <- joined|>
    filter(which_file==input$file, 
           sheet==input$sheet
           )|>
    pull(joined)
  
  temp[[1]]|>
    #filter(year>input$year)|>
    filter(year>=year(today()))|>
    group_by(variable)|>
    mutate(original=if_else(min(abs(original_value)) < input$minimum, "small", "large"),
           change=if_else(original=="small", 
                          new_value-original_value, 
                          (new_value-original_value)/abs(original_value)*100))|>
    select(-new_value, -original_value, -original)|>
    pivot_wider(names_from = "year", values_from = "change")|>
    column_to_rownames("variable")|>
    remove_empty("rows")
})

reactive_show_labels <- reactive({
  if_else(nrow(reactive_tbbl())<80, TRUE, FALSE)
})
```
 
Column {data-width=600}
-------------------------------------
    
### `r renderUI({ paste(input$file, input$sheet, sep=": ")})`
    
```{r}
js <- "
function(el, x) {
  el.on('plotly_click', function(d) {
    var point = d.points[0];
    Shiny.setInputValue('point', {y: point.y});
  });
}"

plotly::renderPlotly({
    heatmaply::heatmaply(reactive_tbbl(), 
                         dendrogram=FALSE, 
                         key.title="Change between cuts", 
                         fontsize_row = 8, 
                         showticklabels = c(TRUE, reactive_show_labels()))|>
    plotly::config(displayModeBar = FALSE)|>
    onRender(js)
})
```

Column {data-width=400}
-------------------------------------

### Click on heatmap to show the row

```{r}
plotly::renderPlotly({
  req(input$point)
  reactive_tbbl()|>
    rownames_to_column("thing")|>
    pivot_longer(cols = -thing)|>
    filter(thing==rev(rownames(reactive_tbbl()))[input$point$y])|>
    ggplot(aes(value, name))+
    geom_col()
})
```



Comparison with internal employment forecast
=====================================  

Column
-------------------------------------
    
### % change from internal forecast to stokes forecast
    
```{r}
internal_vs_stokes|>
  mutate(stokes_percent_change=round((stokes_cut/internal-1)*100,2))|>
  select(industry, name, stokes_percent_change)|>
  filter(name>2023)|>
  pivot_wider(id_cols = industry, names_from = "name", values_from = "stokes_percent_change")|>
  column_to_rownames("industry")|>
  heatmaply::heatmaply(dendrogram=FALSE, key.title="% change", showticklabels = c(TRUE, FALSE))
```
 
Column
------------------------------------- 
 
### First five year CAGRs
    
```{r}
cagr_plot(cagrs, internal_cagr_ffy, stokes_cagr_ffy)
``` 


### Second five year CAGRs
    
```{r}
cagr_plot(cagrs, internal_cagr_sfy, stokes_cagr_sfy)
```

Column
------------------------------------- 

### Ten year CAGRs

```{r}
cagr_plot(cagrs, internal_cagr_ty, stokes_cagr_ty)
```

### A high standard deviation may indicate weird forecast shape...

```{r}
plt <- internal_vs_stokes|>
  group_by(industry)|>
  summarise(mean_growth=mean(stokes_growth, na.rm = TRUE),
            sd_growth=sd(stokes_growth, na.rm=TRUE)
            )|>
  ggplot(aes(mean_growth, sd_growth, text=industry))+
  geom_point(alpha=.5)+
  labs(x="Mean annual growth rate",
       y="Standard deviation growth rate")
plotly::ggplotly(plt, tooltip = "text")
```

Relative to LFS Data
===================================== 

Inputs {.sidebar}
-------------------------------------

```{r}
selectInput(
  "industry",
  "Select an industry",
  unique(internal_vs_stokes$industry),
  selected = unique(internal_vs_stokes$industry)[1]
)

reactive_tbbl2 <- reactive({
 internal_vs_stokes|>
    mutate(date=ymd(paste(name,"06","15", sep="-")))|>
    filter(date>today())|>
    select(-name, -stokes_growth)|>
    pivot_longer(cols=c(stokes_cut, internal), names_to = "series", values_to = "value")|>
    full_join(lfs_data)|>
    filter(industry==input$industry)
})

reactive_lfs_last_year <- reactive({
  lfs_data|>
    filter(industry==input$industry)|>
    group_by(industry)|>
    slice_max(date, n=12)|>
    mutate(value=mean(value))|>
    mutate(series="Last 12 month average of LFS")
})


```

Column
------------------------------------- 

### `r renderUI({input$industry})`


```{r, fig.retina=2, fig.width=18, fig.height=8}
plotly::renderPlotly({
  plt <- ggplot(reactive_tbbl2(), aes(date, value, colour=series))+
    geom_line()+
    geom_line(data=reactive_lfs_last_year())+
    scale_y_continuous(trans="log10", labels = scales::comma)+
    scale_colour_brewer(palette = "Dark2")+
    labs(x=NULL,
         y=NULL,
         colour=NULL)
  plotly::ggplotly(plt)
})
```


Mapping from LMO industries to Stokes industries
===================================== 

```{r}
DT::datatable(mapping, rownames = FALSE, options = list(paging = FALSE))
```







